 ---
title: "p8105_hw2_xl2934"
author: "Xiaoyang Li"
date: "2019/9/27"
output: github_document
---
```{r set up, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(tidyverse)
library(readxl)
```


# Problem 1

## Read and clean the Mr. Trash Wheel sheet

```{r}
trashwheel = 
  read_excel(
    path = "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
    range = "A2:N52",
    sheet = "Professor Trash Wheel"
    ) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(sports_balls))
trashwheel
```

## Read and clean precipitation data from 2017 and 2018
```{r}
prec_2018 = 
  read_excel(
    path = "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
    sheet = "2018 Precipitation",
    range = "A2:B14") %>% 
  mutate(
    year = "2018")

prec_2017 = 
  read_excel(
    path = "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
    sheet = "2017 Precipitation",
    range = "A2:B14") %>% 
  mutate(
    year = "2017")

prec = bind_rows(prec_2017, prec_2018) %>% 
  janitor::clean_names() 
prec = mutate(prec, month = month.name[pull(prec, month)])
```

## Write about data

For `trashwheel`, there are `r nrow(trashwheel)` observations after  omiting non-data rows and non dumpster-specific rows. The median number of sports balls in a dumpster in 2017 is `r filter(trashwheel, year == 2017) %>% pull(sports_balls) %>% median()`. In this dataset, `dumpster` is key variable as far as I am concerned. It contain the unique value in this dataset so that it can help mark every observation. Also, it map to unique date, which make it possible to merge this dataset to others.

For `prec`, there are `r nrow(prec)` observations after  omiting NAs. The total precipitation in 2018 is `r filter(prec, year == 2018) %>% pull(total) %>% sum()`. The key variable in this dataset is `total`, which represent the preciptation in every month.


# Problem 2
## clean the data in pols-month.csv.
```{r}
pols_month = read_csv(file = "./data/pols-month.csv") %>% 
  separate(mon, c("year", "month", "day")) %>% 
  mutate(
    year = as.numeric(year), 
    month = as.numeric(month)) 
  
pols_month = mutate(pols_month, month = month.name[pull(pols_month, month)]) %>% 
  pivot_longer( 
    c("prez_gop", "prez_dem"), 
    names_to = "president", 
    names_prefix = "prez_") %>% 
  select(-value, -day)
pols_month
```
## clean the data in snp.csv 
```{r}
snp = read_csv(file = "./data/snp.csv") %>% 
  separate(date, c("month", "day", "year")) %>% 
  mutate(
    month = as.numeric(month), 
    day = as.numeric(day), 
    year = as.numeric(year)
    ) %>% 
  arrange(year, month) %>% 
  select( year, month, everything(), -day)

snp = mutate(snp, month = month.name[pull(snp, month)])
snp
```

## join the datasets
```{r}
pol_snp = left_join(pols_month, snp, by = c("year", "month"))

unemployment = read_csv(file = "./data/unemployment.csv") %>% 
  rename(
    "year" = "Year",
    "January" = "Jan",
    "February" = "Feb",
    "March" = "Mar",
    "April" = "Apr",
    "June" = "Jun",
    "July" = "Jul",
    "August" = "Aug",
    "September" = "Sep",
    "October" = "Oct",
    "November" = "Nov",
    "December" = "Dec"
    ) %>% 
  pivot_longer(
    January:December,
    names_to = "month",
    values_to = "umemploy_%"
  ) 
  
pol_snp_unemploy = left_join(pol_snp, unemployment, by = c("year", "month"))
```

## write about datasets
For `pols_month`, It contain `r colnames(pols_month)`.
For `snp`, It contain`r colnames(snp)`.
For `unemployment`, It contain`r colnames(unemployment)`.
For the resulting dataset, `pol_snp_unemploy`, there are `r nrow(pol_snp_unemploy)` observations and `r ncol(pol_snp_unemploy)` variables. This dataset record from `r min(pull(pol_snp_unemploy, year))` to `r max(pull(pol_snp_unemploy, year))`. The key variables used to merge different dataset is `year` and `month`.

# Problem 3
## Load and tidy the data
```{r}
names = read_csv(file = "./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  distinct() %>% 
  mutate(ethnicity = recode(ethnicity, 
                            "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER", 
                            "BLACK NON HISP" = "BLACK NON HISPANIC", 
                            "WHITE NON HISP" = "WHITE NON HISPANIC"),
         childs_first_name = str_to_title(childs_first_name))

```

## produce new tables of "Olivia" and most popular male name
```{r}
pop_olivia = filter(names, childs_first_name == "Olivia", gender == "FEMALE") %>% 
  select(year_of_birth, ethnicity, rank) %>% 
  pivot_wider(
    names_from = "ethnicity",
    values_from = "rank"
  )

pop_male = filter(names, gender == "MALE", rank == 1) %>% 
  select(year_of_birth, ethnicity, childs_first_name)%>% 
  pivot_wider(
    names_from = "ethnicity",
    values_from = "childs_first_name"
  )

```

## produce a scatter plot for male, white non-hispanic children born in 2016
```{r fig.height=35}

scatter_samp = filter(names, 
                      gender == "MALE", 
                      ethnicity == "WHITE NON HISPANIC", 
                      year_of_birth == 2016) %>% 
  select(childs_first_name, rank)

ggplot(scatter_samp, aes(x = rank, y = childs_first_name)) + 
  geom_point()


```

