---
title: "p8105_hw2_xl2934"
author: "Xiaoyang Li"
date: "2019/9/27"
output: github_document
---
```{r set up, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(tidyverse)
library(readxl)
```


# Problem 1

## Read and clean the Mr. Trash Wheel sheet

```{r}
trashwheel = 
  read_excel(
    path = "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
    range = "A2:N52",
    sheet = "Professor Trash Wheel"
    ) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(sports_balls))
trashwheel
```

## Read and clean precipitation data from 2017 and 2018
```{r}
prec_2018 = 
  read_excel(
    path = "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
    sheet = "2018 Precipitation",
    range = "A2:B14") %>% 
  select(-Month) %>% 
  mutate(
    year = "2018", 
    month = month.name)

prec_2017 = 
  read_excel(
    path = "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
    sheet = "2017 Precipitation",
    range = "A2:B14") %>% 
  select(-Month) %>% 
  mutate(
    year = "2017",
    month = month.name)

prec = bind_rows(prec_2017, prec_2018) %>% 
  janitor::clean_names() %>% 
  drop_na(total)

```

## Write about data

For `trashwheel`, there are `r nrow(trashwheel)` observations after  omiting non-data rows and non dumpster-specific rows. 

For `prec`, there are `r nrow(prec)` observations after  omiting NAs. The total precipitation in 2018 is `r filter(prec, year == 2018) %>% pull(total) %>% sum()`. The median number of sports balls in a dumpster in 2017 is `r filter(trashwheel, year == 2017) %>% pull(sports_balls) %>% median()`.

But how to choose key variables???

